---
title: "Clustering on Cheese dataset"
author: "Roza Gruter"
date: "9/23/2020"
output:
  html_document:
    theme: yeti
---

##### **Tags:** #Clustering #Exercice #HAC #K-means

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r echo=FALSE}
library(GGally)
library(kableExtra)
library(ggplot2, ggdendro)
```

***

### Introduction
_Fromage_ dataset is a dataset containing 29 fromages and their nutritional information.
The goal of this exercice is to determine groups of homogenous cheeses based on the nutritional information.
II'll use two clustering algorithms: Hierarchical Agglomerative Clustering (HAC) and K-means.

_This exercice is taken from the [teaching page of Marie Chavent](http://www.math.u-bordeaux.fr/~mchave100p/teaching/) (University of Bordeaux)._

### Importing dataset. Descriptive Statistics

```{r}
dta <- read.table("DATA/fromage.txt", header=TRUE, row.names=1)
```

```{r}
head(dta) %>%
  kbl(caption = "Extract of Cheese dataset") %>%
  kable_styling(font_size = 10) %>%
  kable_minimal()
```

```{r echo=FALSE}
summary(dta) %>%
  kbl(caption = "Basic statistics") %>%
  kable_styling(font_size = 10) %>%
  kable_minimal()
```

```{r echo=FALSE}
std_dev <- round(apply(dta, 2, sd), 2)

std_dev %>%
  kbl(caption = "Quick glance at standard deviation for each variable", col.names = c("Standard deviation"), align="l") %>%
  kable_styling(font_size = 10) 
```
- Range of variable values differ strongly
- Standard deviation ranges from approximatively 7 to 108 units
- No missing values for any of the variables

Given the variations of the variables standardization is necessary.
\
\

#### Correlations
```{r fig.height=7, fig.width=10, message=FALSE}
ggpairs(dta)
```
There seems to be strong links between some variables. For example we have a strong positive correlation between _calories_ and _lipides_, or _lipides_ and _cholesterol_, or a negative correlation between _calcium_ and _folates_.
\
\

#### Data standardization
```{r}
n <- nrow(dta)
dta_standard <- as.data.frame(scale(dta, center=TRUE,scale=TRUE)*sqrt(n/(n-1)))

round(dta_standard[1:4,1:4], 2) %>%
  kbl(caption = "Quick glance at standardized values", align="l") %>%
  kable_styling(font_size = 10)
```
\

### Hierarchical Agglomerative Clustering (HAC) using Ward algorithm
```{r}
d <- dist(dta_standard)
tree <- hclust(d^2/(2*n), method="ward.D")
```
\
\

#### Choosing number of classes
```{r fig.width=10, fig.height=6}
par(mfrow=c(1,2))
barplot(sort(tree$height, decreasing = TRUE), main="Heights of the tree")
plot(tree, hang=-1, main="HAC Ward", sub="", xlab="")
```

Adding dividers to decide the number of classes to keep. Judgind by the heights of the tree branches (barplot), important loss of information happens between 4 and 5 classes, so 5 is retained.

```{r fig.width=10, fig.height=6}
par(mfrow=c(1,2))
barplot(sort(tree$height, decreasing = TRUE), main="Heights of the tree")
abline(h = 0.4, col="red")

plot(tree, hang=-1, main="HAC Ward", sub="", xlab="")
rect.hclust(tree, k=5)
```
\
\

#### Cutting the tree obtained wth hclust()
A new variable _part_ is create to store cheeses and assignes class.
```{r}
K <- 5 # number of classes
part <- cutree(tree,k=K)
head(part) %>%
  kbl(caption = "First few cheeses with assigned class", col.names = c("Assigned class"), align="l") %>%
  kable_styling(font_size = 10) 
```

### Interpretation of clustering using PCA
```{r}
library(FactoMineR)
part <- as.factor(part)
levels(part) <- paste("cluster",1:K,sep=" ")
res <- PCA(data.frame(part, dta), quali.sup=1, graph=FALSE) # ACP normée
res$eig # bonne représentation sur le plan factoriel 1-2 (76% d'information retrouvée)
```

```{r fig.width=7, fig.height=5}
par(mfrow=c(1,2))
plot(res,habillage=1,invisible="quali",title="") # graphique des fromages sur le plan 1-2
plot(res,choix="var",title="") # cercle des corrélations sur le plan 1-2
```


### Interpretation of clustering result using quantitative variables
```{r}
des <- catdes(data.frame(part,dta), num.var=1)
```

```{r}
des$quanti.var
```


#### Cluster 5
```{r}
print(des$quanti$`cluster 5`[,1:5],digits=2) # cluster5
```




