---
title: "Web Scraping: how to get data from the web"
author: "Roza Gruter"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document:
  theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r, include=FALSE}
library(rvest)
library(stringr)
library(dplyr)
library(xml2)
library(kableExtra)
```

### Objective
I'm working on an analysis of dairy products (yogurts - regular and those for babies). I'm using web scraping to get the relevant data directly from the net.
Here I'm scraping Auchan's internet store for information on each product: name, producer, price, ingredients and nutritional information.
\
I'm assuming some basic knowledge of html, necessary to find the information we're looking for.
\

##### Introduction to web scraping with rvest

1. First, I need a URL to scrape
2. Then I read the content of the URL into a variable with **read_html(_URL_)**
3. Once the content is stored, I look for the information which is contained in html nodes. I use **html_nodes(nodeName)** function to find the nodeName I'm interested in (I need to find it beforehand, this is where basic knowledge of html comes in handy). then I use **html_attr(attributeName)** function to read the content of the attributeName
4. Then, I'll be using two other functions:
    + **html_attr(attributeName)** to read the content of the attributeName - to get URLs of all products from a result page
    + **html_text()** to read the content of a given node
\

#### Before the fun stuff starts

Creation of URL variables, to go through multiple result pages
```{r}
url_baby_auch <- "https://www.auchan.fr/bebe/desserts-gouters-de-bebe/desserts-lactes/ca-n080301?page="
url_adults_auch <- "https://www.auchan.fr/produits-laitiers-oeufs-fromages/yaourts-fromages-blancs/ca-n0102?page="
base_url_auch <- "https://www.auchan.fr"
```
\

#### Getting URLs for each page of results for baby products
The results are on 2 pages. First I create URLs for each page of results
```{r}
# variable to store URLs for each page of results
pages_auch_baby <- data.frame("URL"=NULL)

# to get the URLs of the 2 pages of results
for (i in 1:2) {
  temp <- data.frame("URL"=paste(url_baby_auch, i, sep=""))
  # print(url_adults_temp)
  pages_auch_baby <- rbind(pages_auch_baby, temp)
}
```
\

#### Getting URLs for each product for babies
I take the website with the list of all products for babies and I read _href_ attribut to get links to each product's page
```{r}
# variable to store URLs for each product
href_baby_auch <- data.frame("URL"=NULL)

for (i in dim(pages_auch_baby)[1]) {
  list_href_baby_auch <- read_html(as.character(pages_auch_baby[i,1])) %>%
    html_nodes('.product-thumbnail__details-wrapper') %>%
    html_attr("href")
  
  temp <- data.frame("URL"=list_href_baby_auch)
  href_baby_auch <- rbind(href_baby_auch, temp)
}
```
\

#### Getting URLs for each page of results for regular products
The results are on multiple pages. We need to first create URLs for each page of results
```{r}
# variable to store URLs for each page of results
pages_auch_adults <- data.frame("URL"=NULL)

# to get the URLs of all pages of results
for (i in 1:13) {
  temp <- data.frame("URL"=paste(url_adults_auch, i, sep=""))
  pages_auch_adults <- rbind(pages_auch_adults, temp)
}
```
\
Then I scrape each page for _href_ attribute to get urls of each product
```{r}
# variable to store URLs for each product
href_adults_auch <- data.frame("URL"=NULL)

# to get URLs of all results from one page
for (i in 1:dim(pages_auch_adults)[1]) {
  list_href_adults_auch <- read_html(as.character(pages_auch_adults[i,1])) %>%
    html_nodes('.product-thumbnail__details-wrapper') %>%
    html_attr("href")
  
  temp <- data.frame("URL"=list_href_adults_auch)
  
  href_adults_auch <- rbind(href_adults_auch, temp)
}
```
\

#### Combine data frames of page URLs of regular and baby products
```{r}
pages_auch_adults <- data.frame(type="adult", URL=pages_auch_adults$URL)
pages_auch_baby <- data.frame(type="baby", URL=pages_auch_baby$URL)
details_auch <- rbind(pages_auch_adults, pages_auch_baby)
```
\

#### Combine data frames of product URLs of regular and baby products
```{r}
href_adults_auch <- data.frame(type="adult", URL=paste(base_url_auch, href_adults_auch$URL, sep=""))
href_baby_auch <- data.frame(type="baby", URL=paste(base_url_auch, href_baby_auch$URL, sep=""))
details_auch <- rbind(href_adults_auch, href_baby_auch)
```
\

#### Scrape each product's website
I want to get nutrition info for each product and to combine all information into one data frame
```{r}
# placeholder for complete dataframe
auchan_raw <- data.frame(shop="Auchan", URL=NA, type=NA, name=NA, producer=NA, ingredients=NA, stringsAsFactors=FALSE)

for (i in 1:dim(details_auch)[1]) {
  html <- read_html(as.character(details_auch[i,2]))
  
  name <- html_nodes(html,'.site-breadcrumb__title') %>%
    html_text()
  
  producer <- html_nodes(html,'.product-description__feature-wrapper:nth-child(2) .product-description__feature-title+ .product-description__feature-group-wrapper .product-description__feature-value') %>%
    html_text()
  
  ingredients <- html_nodes(html,'.product-description__feature-wrapper:nth-child(1) .product-description__feature-value') %>%
    html_text()
  
  nutrition_head <- html_nodes(html,'.nutritional__head') %>%
    html_text()
  nutrition_cell <- html_nodes(html,'.nutritional__cell') %>%
    html_text()
  
  if (length(nutrition_head) != 0) { # if we have the nutrition information for a given product
    # calculations to get column and cell number
    total_cells <- length(nutrition_cell) # how many cells do we have for a given product
    nb_vars <- total_cells/length(nutrition_head) # number of variables (no division on daily value and per 100g value)
    nb_cols <- 3
    nb_total <- nb_vars*nb_cols
    
    # checking if I have all the information, or do I need to add cells
    if (nb_total==total_cells) {
      # creating a temporary variables to hold updated variable names
      tmp <- as.vector(sapply(nutrition_cell[1:nb_vars], paste, "(ap100g/ml)", sep=" "))
      temp <- as.vector(sapply(nutrition_cell[1:nb_vars], paste, "(ap_jour_%)", sep=" "))
      
      # modifying nutrition_cell to add new variable names
      # index for values of (ap100g/ml)
      b <- nb_vars+1
      c <- nb_vars*2
      d <- c+1
      # updated nutrition_cell vector
      nutrition_cell <- c(tmp, temp, nutrition_cell[b:c], nutrition_cell[d:nb_total])
    } else {
      tp <- rep(NA, times=nb_total-total_cells) # NA vector to be added to the existing cells
      nutrition_cell <- c(nutrition_cell, tp) # adding NAs to the existing cells
      # creating a temporary variables to hold updated variable names
      tmp <- as.vector(sapply(nutrition_cell[1:nb_vars], paste, "(ap100g/ml)", sep=" "))
      temp <- as.vector(sapply(nutrition_cell[1:nb_vars], paste, "(ap_jour_%)", sep=" "))
      
      # modifying nutrition_cell to add new variable names
      # index for values of (ap100g/ml)
      b <- nb_vars+1
      c <- nb_vars*2
      d <- c+1
      # updated nutrition_cell vector
      nutrition_cell <- c(tmp, temp, nutrition_cell[b:c], nutrition_cell[d:nb_total])
    }
    
    # add an entry to the main data frame data_auchan
    zero <- rep(0, times=dim(auchan_raw)[2]-4)
    auchan_raw <- rbind(auchan_raw, c("Auchan", as.character(details_auch[i,2]), as.character(details_auch[i,1]), name, producer, ingredients[1], zero))
    
    # create a dataframe variable/value for nutrition info
    x <- length(nutrition_cell)/2
    nutrition_df <- data.frame(t(nutrition_cell[1:x]), stringsAsFactors=FALSE)
    colnames(nutrition_df) <- nutrition_cell[1:x]
    tps <- data.frame(t(nutrition_cell[(x+1):length(nutrition_cell)]), stringsAsFactors=FALSE)
    colnames(tps) <- nutrition_cell[1:x]
    nutrition_df <- rbind(nutrition_df, tps)
    nutrition_df <- nutrition_df[-1,]
  }

  # add nutrition information to the data_auchan dataframe
  g <- dim(auchan_raw)[1]
  for (i in 1:length(nutrition_df)) {
    if (!(colnames(nutrition_df)[i] %in% colnames(auchan_raw))) { # if variable doesn't exist in the dataframe
      auchan_raw[g,length(auchan_raw)+1] <- nutrition_df[i] # I add a new column
    } else {
      id <- which(colnames(auchan_raw)==colnames(nutrition_df)[i])
      auchan_raw[g,id] <- nutrition_df[[i]]
    }
  }
}

auchan_raw <- auchan_raw[-1,-1]
```
\

#### A quick look at the raw data
```{r}
head(auchan_raw[,c(2, 3, 5, 6)], 5) %>%
  kbl(caption = "Dairy prodycts raw data - extract") %>%
  kable_minimal()
```
\

#### Save the raw file
To finish I save the raw data into a file.
```{r}
write.csv2(auchan_raw, file="DATA/auchan_raw.csv", sep=";", row.names=FALSE)
```
\

###### Thanks for reading!!
\
\