---
title: "Data Cleaning: putting order in raw dataset"
author: "Roza Gruter"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    theme: yeti
    toc: true
    toc_depth: 5
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(stringr)
library(kableExtra)
library(ggplot2)
library(naniar)
```

```{r include=FALSE}
# importing dataset
data <- read.csv2("DATA/auchan_raw.csv", sep=",")
```
\

### The data
This paper is Part 2 of dairy products analysis project. Part 1 (web scraping) can be checked [here](https://rozagruter.github.io/content/projects/1_WebScraping-yogurts.html).
\
\

#### Data at a glance
A quick look at a selection of four variables from raw dataset
```{r echo=FALSE}
head(data[,c(4,7:9)])
```
\

#### Structure
```{r echo=FALSE}
str(data[7:10])
```

There are several problems that we can spot immediately:

- first column is just an index, so it can be deleted
- variable names are a mess, we will have to rename them
- most variables are factors while they should be numeric
- ordering of the variables is a mess as well
\

### Cleaning the data

#### 1. Deleting first column
First column of the dataset is an old index column, it is deleted.
```{r}
data$X <- NULL
```
\

#### 2. Transforming characters into numbers
All the numeric data have been scraped as strings with a special character present (\n). Here I'm replacing "\n" with nothing (""), then I remove all the leading and trainling spaces, and I transform the result into a number.
```{r}
for (i in 6:length(data)) {
  if (!is.numeric(data[,i])) {
    data[i] <- gsub("[\n]", "", data[,i]) %>%
      trimws() %>%
      as.numeric()
  }
}
```
\

Quick look at the dataset now that variables are correctly displayed as numeric
```{r}
str(data[6:10])
```
\

#### 3. Renaming columns
I start by putting existing column names into a vector to work on a separate object. The data frame will only be updated once all the modifications are validated.
```{r}
c.names <- colnames(data)
```
\

##### First part: I'm taking care of formatting dosage information
```{r}
c.names <- str_replace(c.names, "..en.g...ap_jour_..", ".ppj") %>%
  str_replace(".en.Kj..ap_jour_..", ".pkjpj") %>%
  str_replace(".en.Kj..ap100g.ml.", ".kjp100") %>%
  str_replace(".en.Kcal..ap100g.ml.", ".kcalp100") %>%
  str_replace(".en.Kcal..ap_jour_..", ".pkcalpj") %>%
  str_replace("..en.g...ap100g.ml.", ".gp100") %>%
  str_replace("..en.µg...ap100g.ml.", ".µgp100") %>%
  str_replace("..en.µg...ap_jour_..", ".ppj") %>%
  str_replace("..en.mg...ap100g.ml.", ".mgp100") %>%
  str_replace("..en.mg...ap_jour_..", ".ppj")
```
\

##### Second part: I'm taking care of french caracters and I'm shortening variable names
```{r}
c.names <- str_to_lower(c.names) %>%
  str_replace_all(c("é"="e", "è"="e", " "=".", "µg"="mcg",
                  "valeurs.energetiques"="val.energ",
                  "matieres.grasses"="mat.grasses",
                  "fibres.alimentaires"="f.aliment",
                  "acide.folique"="a.folique",
                  "acide.pantothenique"="a.pantothenique"))
```
\

##### Third part: I'm replacing old column names with new ones
```{r}
colnames(data) <- c.names
```
\

#### 4. Ordering columns
I keep the variables like _type_ and _name_ at the beginning, then I reorder the rest putting the value in respective units (grammes, miligrammes etc.) first, followed by percentage of daily recommended intake.
\

I start by manually ordering 21 first variables. I store them in _sorted.names_ object
```{r}
sorted.names <- c(c.names[1:5], c.names[7], c.names[15], c.names[6], c.names[14], c.names[8:9], c.names[16:17], c.names[10:11], c.names[18:19], c.names[12], c.names[20], c.names[13], c.names[21])
```
\

Then I create a temporary object _temp_ and store in it all the remaining variables that I sort first. I continue by adding those variables to the _sorted.names_ object.
```{r}
temp <- sort(c.names[!(c.names %in% sorted.names)])
sorted.names <- c(sorted.names, temp)
```
\

The last step is to reorder columns in the dataset
```{r}
data <- data[, sorted.names]
```
\

### Result
Here is an extract of cleaned dataset:
```{r}
head(data[,c(4,6:9)]) %>%
  kbl(caption = "Extract of cleaned dataset") %>%
  kable_minimal(font_size = 12, bootstrap_options=c("stripped", "hover"), full_width=F, position="left")
```
\

#### Note on missing variables
I completely left out missing values management in this section. There is a very high number of missing values in this dataset. 
```{r fig.width=8}
nas <- miss_var_summary(data) %>%
  select("variable", "n_miss")

ggplot(nas) +
  aes(y=n_miss, x=variable) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust=1))
```
\

Overall `r sum(nas$n_miss>0)` variables have missing values, out of which `r sum(nas$n_miss>30)` have more than 30 missing values. I would like to find a way of replacing them instead of dropping observations or variables. In order to do so I'll work on missing values for each variable separately while performing exploratory analysis.
\

### Save the cleaned file
To finish I save the cleaned data into a file.
```{r}
write.csv2(data, file="DATA/auchan_cleaned.csv", row.names=FALSE)
```
\

##### Thanks for reading!!
\
\
